---
title: "Aides Drug"
author: "Matthew Ramirez"
date: "Fall 2025"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the females, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

I noticed that the data is using the Categorical Categorical data, and in order to solve this problem the Categorical Categorical helper file is needed.

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
Females <- filter(triUmp, gender=="F")
```

```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")

```

What I observed in the bar chart, I realized the people who took the drug of the 58% got aids, and the 40% of the people who took it who got aids were less than those who took the drug. The drug did the opposite rather than prevent people from getting aids.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)
colPerc(table1)
```

After looking over the numerical results I noticed that graphical results corresponds correctly, proving that the drug didn't work and made it worse as 58.82% of people got aids.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

After looking over the odds ratio it is 1.5 times more likely to get aids from the medication rather than the placebo itself. The p-values itself from the 3 different tests (Pearson's Chi-squared, Chi-squared test, and fishers test) gave a low p-value showing no significant difference in medication and placebo ultimately making the medication more dangerous. The p-value 7.686e-05, 1e-04, and 6.596e-05 prove the data.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

Based off the graphical result the medication caused more people to gain aids rather than the placebo. The numerical results corresponded with the graphical results of how 58.82% of the people who took the medication got aids in an exact amount. The inferential results then took the p-values from the 3 tests and estimated that there's no significant difference in the medication and placebo ultimately concluding that the odds of the medication giving people aids is 1.5 times more likely than the placebo effect.
